---
title: "Journal_exercise"
author: "Amieroh Abrahams"
date: "06 August 2018"
output: html_document
---
First I need to find, install and load various packages. These packages will be available on CRAN and can be accessed and installed in the usual way.

```{r}
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(ggthemes)
library(readr)
```

The class collected data from three types of journals. The journal of biogeography, journal of ecography and the journal of ecology. The first 5 publications of each of the years were selected. The class was divided into four groups. Each group selected one decade ranging from 1970, 1980, 1990, 2000. Each group recorded the number of authors, the size of the study and the year of publication for their decade.

Here i load in the dataset. Continuous data

```{r}
# Journal_data <- read_csv("Data/Journal_data.csv")
JD_2000 <- read_csv("Data_Journals/JD_2000.csv")
```

Plotting a bar graph for the year 1970s

```{r}
plot1 <- ggplot(Biogeography_1970s, aes(x = year, y = No_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Biogeography 1970's") +
  theme_pubclean() + scale_color_few() +
  scale_fill_few()

plot2 <- ggplot(Ecology_1970s, aes(x = year, y = No_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecology 1970's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

plot3 <- ggplot(Ecography_1970s, aes(x = year, y = No_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecography 1970's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

combined_1970s <- ggarrange(plot1, plot2, plot3,
          ncol = 2, nrow = 2)
combined_1970s
```

Loading in the data for the year 1980

```{r}
JD_1980 <- read_csv("Data/JD_1980.csv")
```

```{r}
JD_biogeography <- JD_1980 %>% 
  filter(journal == "biogeography")

graph1 <- ggplot(JD_biogeography, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Biogeography 1980's") +
  theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_ecology <- JD_1980 %>% 
  filter(journal == "ecology")

graph2 <- ggplot(JD_ecology, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecology 1980's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_ecography <- JD_1980 %>% 
  filter(journal == "ecography")

graph3 <- ggplot(JD_ecography, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecography 1980's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

combined_1980s <- ggarrange(graph1, graph2, graph3,
          ncol = 2, nrow = 2)
combined_1980s
```

Testing for normality

```{r}
# normality_test_1980 <- JD_1980 %>% 
#   group_by(year) %>% 
#   summarise(norm_authors = as.numeric(shapiro.test(no_of_authors)[2]))
# 
# # P value appears to be less than 0.05
# # data is not normally distributed
```

Here i load in the year 1990 dataset

```{r}
JD_1990 <- read_csv("Data/JD_1990.csv") # Excluding matthews data
```

```{r}
JD_bio1990 <- JD_1990 %>% 
  filter(journal == "Biogeography")

gr1 <- ggplot(JD_bio1990, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Biogeography 1990's") +
  theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_eco1990 <- JD_1990 %>% 
  filter(journal == "Ecology")

gr2 <- ggplot(JD_eco1990, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecology 1990's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_ecog1990 <- JD_1990 %>% 
  filter(journal == "Ecography")

gr3 <- ggplot(JD_ecog1990, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecography 1990's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

combined_1990s <- ggarrange(gr1, gr2, gr3,
          ncol = 2, nrow = 2)
combined_1990s
```

Testing for normality

```{r}
# normality_test_1990 <- JD_1990 %>% 
#   group_by(year) %>% 
#  summarise(norm_authors = as.numeric(shapiro.test(no_of_authors)[2]))

# # P value appears to be less than 0.05
# # data appears to be not normally distributed
```

Here i load in the year 2000 dataset

```{r}
JD_2000 <- read_csv("Data/JD_2000.csv")
```

Now i filter and plot each of the journals accordingly

```{r}
JD_bio <- JD_2000 %>% 
  filter(journal == "Biogeography")

g1 <- ggplot(JD_bio, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Biogeography 2000's") +
  theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_eco <- JD_2000 %>% 
  filter(journal == "Ecology")

g2 <- ggplot(JD_eco, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecology 2000's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

JD_ecog <- JD_2000 %>% 
  filter(journal == "Ecography")

g3 <- ggplot(JD_ecography, aes(x = year, y = no_of_authors)) +
  geom_col(aes(fill = study_size)) +
labs(x = "Year", y = "No of authors", colour = "study area") + 
theme(legend.position = "bottom") +
ggtitle("Journal of Ecography 2000's") +
theme_pubclean() + scale_color_few() +
  scale_fill_few()

combined_2000s <- ggarrange(g1, g2, g3,
          ncol = 2, nrow = 2)
combined_2000s
```

Testing for normality

```{r}
#  normality_test_2000 <- JD_2000 %>% 
#    group_by(year) %>% 
#   summarise(norm_authors = as.numeric(shapiro.test(no_of_authors)[2]))
# # 
# # # P value appears to be less than 0.05
# # # data appears to be normally distributed
```

Here i load in the combined dataset

```{r}
Combined_authors <- read_csv("Data/Combined_authors.csv")
```

Here we test to see if the data are normally distributed or not. Here we will make use of the shapiro wil test as this appears to be most accurate when dealing with small datasets. p<0.05 shows that data is not normally distributed.  Normal datasets are also symmetrical. Kurtosis tests can also be done and a kurtosis value between 2 and -2 can also be regarded as a normally distributed dataset.

The data here appear to be normally distributed when comparing the no of authors present for each of the years. This is evident by p>0.05. 

```{r}
normality_test <- Combined_authors %>% 
  group_by(year) %>% 
  summarise(norm_authors = as.numeric(shapiro.test(no_of_authors)[2]))
```

Once we established that the data is normally distributed it is now nessasary to test for homoscedasticity.

```{r}
homo <- Combined_authors %>% 
  group_by(year) %>% 
  summarise(auth_var = var(no_of_authors))
```

Any further analysis requires us to write up an hypothesis in order to determine which tests to use when comparing the combined dataset.


    H0: number of authors does not increase as years go by.
    H1: number of authors increases as years go by.

    H0: The study size has not increased over the past decade.
    H1: The study size has increased over the past decade.

Here we load the ggpubr package. This allows us to access the compare_means() function, which allows us to perform any sort of test that compares sample sets of data and outputs the results as a dataframe. 

Here we are unable to use a ttest because there are more than two factors or groups present within the dataset. So here we do an anova test which looks at the number of authors per year

```{r}
auth.aov <- aov(no_of_authors ~ year, data = Combined_authors)
summary(auth.aov)
  # results show p<0.05...

# if data is not normal... (similar end result though for non-parametric test)
compare_means(no_of_authors ~ year, data = Combined_authors, method = "kruskal.test")
  # results show p<0.05 therefore, 
  # number of authors increase as years go by

# but there are a load of dates, maybe we can condense dates into just 70s, 80s, 90s and 00s?
# then we'll only have four groups to compare instead of all the dates individually?? i think... 

# we can also do an ANOVA for number of authors by journal
# however. not sure if this has any significance to us. but I'll do it anyway
jour.aov <- aov(no_of_authors ~ journal, data = Combined_authors)
summary(jour.aov)
  # results show p = 0.111 so there is no significant difference in number of authors by journal

# same result for non-parametric test
compare_means(no_of_authors ~ journal, data = Combined_authors, method = "kruskal.test")

TukeyHSD(jour.aov)
plot(TukeyHSD(jour.aov))
# TBH im deeply confused and i dont know what this means. i am a fool.

# maybe we can do a correlation to see if theres a relationship between number of authors and
# study size (more authors = larger study size?)
  # cor.test(Combined_authors$no_of_authors, Combined_authors$study_size, method = "")
    # error for this too... SAD. Do you think there will be a difference if we change study size to
    # classes 1-5?

# correlation and regression for year and number of authors
cor.test(Combined_authors$year, Combined_authors$no_of_authors, method = "spearman")
  # 36.8% of variance explained (not very strong relationship but positive)
summary(lm(no_of_authors ~ year, data = Combined_authors))
  # 15.3% of variance explained (also not very strong, positive though)

# random EDA stuff
summary(Combined_authors)

ggplot(data = Combined_authors, aes(x = as.factor(year), y = no_of_authors)) +
  geom_col() +
  facet_wrap(~ journal)

# to try:
  # condense dates into their decades... 70s, 80s etc
    # rename study sizes as class numbers (1-5)... this might make the correlation thing work
      # working with mean number of authors per year or decade may also make things easier 
# in the normality test more of the values are less than 0.05. so i dont think the data is normally distributed...
```






























































### Ignore this!!
Now we examine the dataset. Journal data is the entire class dataset combined

```{r}
Journal_data <- as_tibble(Journal_data)
Journal_data
head(Journal_data)
tail(Journal_data)
summary(Journal_data)
Journal_data %>%
  summarise(mean_Authors =mean(No_of_authors, na.rm = TRUE))
```









